---
title: "Visualising Time Series with Calendar Heatmap with ggplot2"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)
```



```{r}
library(tidyverse)
library(ggthemes)
library(tsbox)
library(bbplot)
library(lubridate)
library(timeDate)
library(forcats)
library(forecast)
```

One way to visualize time series for data on a daily basis are calendar heatmaps. GitHub e.g. is using them to show the users activity.

[Image Github Heatmap]

You could use them to show, if a holiday apartment is available or not. In this case we visualize a binary variable. Or you could visualize a numerical varliavle like the
daily rate of all holiday apartments. This way people can easily spot periods with lower rates. 

However, why should you use a calendar heatmap instead of simple line graph?


By using a calendar heatmap you add some additonal extra dimesion into the visualisation like weekday and the week of the month into graph.
This extra information lets you spot patterns for working days or weekends. Years and months are better seperated than in a simple time line plot. 
The calendar heatmap lets you pack a lot of information on little space without losing too much readability.




```{r}
uk.driver.deaths <- Seatbelts %>% ts_data.frame %>% as_tibble %>% spread(id, value)
```
```{r}
uk.driver.deaths %>% head
```


```{r}
deaths.dec <-
  uk.driver.deaths %>% select(time, DriversKilled) %>% ts_ts %>% decompose(type='multiplicative')
petrol.price.dec <-  uk.driver.deaths %>% select(time, PetrolPrice) %>% ts_ts %>% decompose
drivers.dec <-  uk.driver.deaths %>% select(time, drivers) %>% ts_ts %>% decompose
rear.dec <-  uk.driver.deaths %>% select(time, rear) %>% ts_ts %>% decompose
```
```{r}
plot(drivers.dec)
```




```{r}
deaths.trend <- deaths.dec$trend  %>%
  ts_data.frame %>%
  as_tibble %>%
  na.omit %>%
  mutate(var = "deaths") %>%
  spread(var, value) %>% 
  inner_join( uk.driver.deaths %>% select(time, law), by='time') %>% 
  mutate(law=as.factor(law))

deaths.seasonal <- deaths.dec$seasonal  %>%
  ts_data.frame %>%
  as_tibble %>%
  na.omit %>%
  mutate(var = "deaths") %>%
  spread(var, value) %>% 
  inner_join( uk.driver.deaths %>% select(time, law), by='time') %>% 
  mutate(law=as.factor(law))


petrol.price.trend <- petrol.price.dec$trend  %>%
  ts_data.frame %>%
  as_tibble %>%
  na.omit %>%
  mutate(var = "PetrolPrice") %>%
  spread(var, value) %>% 
  inner_join( uk.driver.deaths %>% select(time, law), by='time') %>% 
  mutate(law=as.factor(law))

petrol.price.seasonal <- petrol.price.dec$seasonal  %>%
  ts_data.frame %>%
  as_tibble %>%
  na.omit %>%
  mutate(var = "PetrolPrice") %>%
  spread(var, value) %>% 
  inner_join( uk.driver.deaths %>% select(time, law), by='time') %>% 
  mutate(law=as.factor(law))
```
```{r fig.width=25}
petrol.price.seasonal %>% 
  ggplot(aes(x = time, y = PetrolPrice, color=law)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month") + 
  theme(axis.text.x = element_text(angle = 45)) +
  facet_wrap(~year(time), scales = 'free')
```


```{r fig.width=25}
deaths.seasonal %>% 
  ggplot(aes(x = time, y = deaths, color=law)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month") + 
  theme(axis.text.x = element_text(angle = 45)) +
  facet_wrap(~year(time), scales = 'free')
```
```{r}
uk.driver.deaths %>% 
  filter(time > '1980-11-01' & time < '1982-04-01')
```


```{r}
deaths.trend %>%
  ggplot(aes(x = time, y = deaths, color=law)) +
  geom_line() +
  scale_x_date(date_breaks = '1 year') +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_color_tableau()
```

```{r}
plot(deaths.dec$trend)
```

```{r}
plot(deaths.dec)
```

```{r}
deaths.seasonal <- deaths.dec$seasonal  %>% ts_data.frame %>% as_tibble %>% na.omit %>% mutate(var="deaths") %>% spread(var, value)
```





```{r}
uk.driver.deaths
```
```{r fig.height=10, fig.width=20}
uk.driver.deaths %>% 
  mutate(law=as.factor(law)) %>% 
  ggplot +
  geom_line(aes(x=time, y=DriversKilled, color=law), size=1.5) +
  scale_x_date(date_breaks = "1 year") +
  scale_color_tableau()
```
```{r}
library(anomalize)
uk.driver.deaths.dec <- uk.driver.deaths %>% decompo
```



```{r}
eu.stock.market <- EuStockMarkets %>% 
  ts_data.frame %>%
  spread(key=id, value=value) %>% 
  mutate(year=year(time), month=month(time, label=T), day=day(time), week=week(time), weekday=as.factor(as.POSIXlt(time)$wday), monthweek=as.factor(1 + week - min(week)))
```




```{r}
eu.stock.market %>% head(30)
```

```{r fig.height=10, fig.width=20}
eu.stock.market %>%
  gather(-time,
         -year,
         -month,
         -day,
         -weekday,
         -weekmonth,
         key = instrument,
         value = value) %>%
  mutate(weekmonth = factor(monthweek, levels = c(5, 4, 3, 2, 1))) %>%
  filter(instrument == 'CAC') %>%
  ggplot(aes(x = weekday, y = monthweek)) +
  geom_raster(aes(fill = value)) +
  geom_text(aes(label = day)) +
  scale_x_discrete(breaks = seq(0, 6, 1),
                   labels = c("Su", "Mo", "Tu", "We", "Th", "Fr", "Sa")) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  facet_grid( ~ year ~ month)
```




```{r fig.height=10, fig.width=20}
eu.stock.market %>% 
  gather(-time, key=index, value=value) %>% 
  ggplot +
  geom_line(aes(x=time, y=value, color=index), size=1) +
  scale_color_tableau() +
  bbc_style()
```



